---
title: "week7"
author: "Sushen S Jamwal"
date: "July 2, 2019"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**# Data Visualization**

## install and load the tidyverse library

```{r}
install.packages('tidyverse',repos='http://cran.us.r-project.org')
library(tidyverse)
```

## To check: Where a function (or dataset) comes from?

```{r}
mpg = ggplot2::mpg
```
##Plotting
###The plot shows a negative relationship between engine size (displ) and fuel efficiency (hwy). In other words, cars with big engines use more fuel. ggplot create a coordinates and then you add layers on top of it. For example: geom_point() a layer of points to your plot.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

The generic ggplot template:

#ggplot(data = <DATA>) + 
#  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))


##Editing

Let's do some editing, aesthetic mappings to change the plot by mapping different variables to the geom function aes. For example:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

Manually assign the value of a specific aesthetic

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

```

##Facts
One way to add additional variables is with aesthetics. Another way, particularly useful for categorical variables, is to split your plot into facets, subplots that each display one subset of the data.
To facet your plot by a single variable, use facet_wrap()
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

```

To facet your plot on the combination of two variables, add facet_grid()
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)

```

##Use different geom objects

geom_point() plots points, however same mapping with geom_smooth() plots in smooth line line

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
 
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))

```

Plot multiple geoms on the same plot,
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```

##Statistical transformations

The chart shows that more diamonds are available with high quality cuts than with low quality cuts.
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

##Position adjustments
position = "fill" works like stacking, but makes each set of stacked bars the same height. This makes it easier to compare proportions across groups.

Similarly, you can use position = "dodge"  and "identity" for different positions
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

##Coordinate systems
default coordinate is cartesian buit we can change using different options available, for example coord_flip() will switch the axes

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()

```